---
title: "CS544_HW5_Vu"
author: "Nguyen Vu"
date: "6/16/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r load packages needed}
library(sampling)
library(prob)
```



##Part1) Central Limit Theorem (20 points)
The input data consists of the sequence from 1 to 20 (1:20). Show the following three
plots in a single row.

```{r 1}
#load data
dataIn <- seq(1:20)
```
a) Show the histogram of the densities of this distribution.
b) Using all samples of this data of size 2, show the histogram of the densities of the
sample means.
c) Using all samples of this data of size 5, show the histogram of the densities of the
sample means.

```{r}
# Use sampling to get the probability space of this vector

par(mfrow = c(1, 3))

#set sample size to draw with replacement as 1, 2 and 5. The mean and sd is calculated.
set.seed(1220)
for(s_size in c(1, 2, 5)){
  sample_x <- urnsamples(dataIn, s_size)
  x_bar = apply(sample_x, MARGIN = 1, FUN = mean)
  hist(x_bar, prob = TRUE, main = paste("Sample Means of n = ", s_size), xlab = "sample mean")
  
  cat("Sample Size = ", s_size, "\n",  "# Observations = ", nrow(sample_x), "\n",  " Mean = ", mean(x_bar), "\n"," SD = ", sd(x_bar), "\n", "\n")
}

```
d) Compare of means and standard deviations of the above three distributions.
  The mean for all distributions are the same = 10.5. For the sample sampling of 1 in part a, the probability density is uniform as there is an equal probability of selecting each number in the vector. 
  As the mean is calculated between 2 samples and 5 samples in part b and c, the number of observations increase with the possible combinations of selecting 5 samples for the mean (20 to 190 to 15504 observations).
  As more mean is added to the probability density, the standard deviation is decreased from 5.9 to 3.98 to 2.29 in part a, b and c respectively. The prob density is approaching normal. 

##Part2) Central Limit Theorem (20 points)
The data in the file queries.csv contains the number of queries Google has had each day for a one
year period (365 days). The data file is also available at
http://kalathur.com/cs544/data/queries.csv. Use this link to read the data using read.csv function
when submitting the homework.

```{r}
#Load data and convert to dataFrame
p2_data <- read.csv("http://kalathur.com/cs544/data/queries.csv", stringsAsFactors = FALSE)
p2_data <- as.data.frame(p2_data)

#get name of column
names(p2_data)
```
a) Show the histogram of the distribution of the number of queries. Compute the mean and
standard deviation of the number of queries Google has had per day.
```{r}
#compute the means and Standard Devs of the # of inquiry:

mu_queries = mean(p2_data$queries)
sd_queries = sd(p2_data$queries)

hist(p2_data$queries, xlim = c(2e8, 3e8), probability = TRUE, main = "distribution of inquiry", 
     xlab = "# of inquiry" )
cat("mean of the number of queries at Google =" , mu_queries,"\n",
"standard deviation of the number of queries at Google =" , sd_queries,"\n")

```
b) Draw 1000 samples of this data of size 5, show the histogram of the densities of the sample
means. Compute the mean of the sample means and the standard deviation of the sample means.
c) Draw 1000 samples of this data of size 20, show the histogram of the densities of the sample
means. Compute the mean of the sample means and the standard deviation of the sample means.
```{r}
set.seed(1220)

x_bar <- numeric(1000)
par(mfrow = c(1, 2))
for(s_size in c(5, 20)){
  for(i in 1:length(x_bar)){
    x_bar[i] = mean(sample(p2_data$queries, size = s_size, replace = TRUE))
  }
  hist(x_bar, prob = TRUE, main = paste("Sample Means of n = ", s_size), xlab = "# of inquiry mean", ylim = c(0, 7e-8), breaks = 50)
  
  cat( ifelse(s_size !=20, "2b", "2c"), "Sample Size = ", s_size,  " Mean = ", mean(x_bar), " SD = ", sd(x_bar), "\n")
}
```


d) Compare of means and standard deviations of the above three distributions.

The shape in part a is of the original sample data population. Of the 365 observations, the sample mean is calucalted to be 248514980 searches with the sd = 29202674. This data has a the probability density shape of uniform distribution.
Keeping the sampling frequency constant at 1000 observations in part b and c, and changing the sampling size to 5 and 20 respectively, the distribution takes a normal curve shapes. The distribution resulted from part c where the sampling size is larger (20>5), the distribution moves closer to resembling a normal curve, the sd is also smaller (reduced from 12519732 in part b to 6644319 in part c)


##Part3) Central Limit Theorem – Negative Binomial distribution (20points)
Suppose the input data follows the negative binomial distribution with the
parameters size = 5 and prob = 0.5.
a) Generate 1000 random numbers from this distribution. Show the barplot
with the proportions of the distinct values of this distribution.

```{r}
#n = 1000, x = 5, prob = 0.5
set.seed(1220)
p3_data <- rnbinom(1000, size = 5, prob = 0.5)

table(p3_data)

hist(p3_data, prob=TRUE, main = "n=1000, size = 4")

```
b) With samples sizes of 10, 20, 30, and 40, generate the data for 5000
samples using the same distribution. Show the histograms of the densities
of the sample means. Use a 2 x 2 layout.
```{r}
sSizes <- c(10, 20, 30, 40)
par(mfrow = c(2, 2))
for(s in sSizes){
  p3_dist <- rnbinom(5000, size = s, prob = 0.5)
  hist(p3_dist, prob = TRUE, main = paste("sample size =", s), xlab = "")
  cat("Sample Size = ", s,  " Mean = ", mean(p3_dist),
        " SD = ", sd(p3_dist), " %CV = ",100*sd(p3_dist)/mean(p3_dist), "\n")
}

```
c) Compare of means and standard deviations of the data from a) with the
four sequences generated in b).
```{r}
cat("Sample Size in part a= ",1000,  " Mean = ", mean(p3_data),
        " SD = ", sd(p3_data)," %CV = ", 100*sd(p3_data)/mean(p3_data), "\n")
```
The sample mean is close to that of the Sample size (5, 10, 20, 30 and 40).
The standard deviation decrease as the sample size increase asa a proportion of the mean. We calculate the %CV to decrease from ~64% to ~22% as sample size increase from 5 to 10. 
In part a, the histogram is also very left-skew and the mean would not be a good description of this data.
In part b, as the sample size increase the distribution takes a more normal shape. 

##Part4) Sampling (40 points)
Use the MU284 dataset from the sampling package. Use a sample size of
20 for each of the following.
```{r}
data("MU284")
names(MU284)
N <- nrow(MU284)
n <- 20


popRegTable <- table(MU284$REG)
```
a) Show the sample drawn using simple random sampling without
replacement. Show the frequencies for each region (REG). Show the
percentages of these with respect to the entire dataset.
```{r}
#n = 20
set.seed(1220)
#random draw without replacement
s <- srswor(n, N)

#get rows number from random sampling
rows <- (1:nrow(MU284))[s!=0]
rows <- rep(rows, s[s!=0])

#subset the rows selected
set_4a <- MU284[rows,] 

s4a_table <- tabulate(set_4a$REG)
#  add misssing component and get % of each REG type
(s4a_table / popRegTable) *100

```

b) Show the sample drawn using systematic sampling. Show the
frequencies for each region (REG). Show the percentages of these with
respect to the entire dataset.
```{r}
set.seed(1220)
pick <- inclusionprobabilities(MU284$REG, n)

s4b <- UPsystematic(pick)

set4b <- MU284[s4b!=0,]
s4b_table <- table(set4b$REG)


#Percentage of each REG category with respect to the population
(s4b_table / popRegTable) *100

```

c) Calculate the inclusion probabilities using the S82 variable. Using these
values, show the sample drawn using systematic sampling. Show the
frequencies for each region (REG). Show the percentages of these with
respect to the entire dataset.

```{r}
set.seed(1220)
pick <- inclusionprobabilities(MU284$S82, n)

s4c <- UPsystematic(pick)

set4c <- MU284[s4c!=0,]
s4c_table <- tabulate(set4c$REG)

#Percentage of each REG category with respect to the population
(s4c_table / popRegTable) *100
```

d) Order the data using the REG variable. Draw a stratified sample using
proportional sizes based on the REG variable. Show the frequencies for
each region (REG). Show the percentages of these with respect to the
entire dataset.
```{r}
set.seed(1220)
#order data
orderedData <- MU284[order(MU284$REG),]

freq <- table(orderedData$REG)
stratsize <- ceiling(n*(freq/sum(freq)))

stra1 <- strata(MU284, stratanames = c("REG"), size = stratsize, method = "srswor", description = FALSE)

stra1_Sample <- getdata(MU284, stra1)
#frequency of each region:
table(stra1_Sample$REG)


s4d_table <- tabulate(stra1_Sample$REG)
#Percentage of each REG category with respect to the population:
(s4d_table / popRegTable) *100



```

e) Compare the means of RMT85 variable for these four samples with the
entire data.

```{r}
rmtmean_a <- mean(set_4a$RMT85)
rmtmean_b <- mean(set4b$RMT85)
rmtmean_c <- mean(set4c$RMT85)
rmtmean_d <- mean(stra1_Sample$RMT85)
rmt_MU <- mean(MU284$RMT85)

cat("The mean of RMT85 variable for these four samples is as follows: \n", 
    "simple random sampling-4a: ",rmtmean_a, "\n",
    "systematic sampling using REG freq-4b:",rmtmean_b,"\n",
    "systematic sampling using the S82 variable-4c",rmtmean_c,"\n",
    "stratify sampling by REG-4d:",rmtmean_d,"\n",
    "Population average =", rmt_MU)


```

Looking at the means of RMT85 from the previous four samples, the systematic sampling for each region REG looks to be the closest to the population mean. 
The systematic sampling using the S82 variable seems to be alot further away from the average.
This may indicate that alot of the high values for RMT85 is embedded within the highest probability of the S82 variables. 
The stratify by REG mean and the simple random sampling mean are very close to each other, this may indicate that the population mean may be skewed by having collected a large number of S82 where the values of RMT85 is also higher. 
Looking at table(MU284$S82), we can see that a large frequency of S82 is collected between 41 and 51. This may also be where RTM85 has a higher value. 