---
title: "CS544_HW6_Vu"
author: "Nguyen Vu"
date: "6/20/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part1) Strings (60 points)
Use the stringr functions for the following:
Initialize the vector of words from Lincoln’s Gettysburg address with the
following code:
file <- "http://kalathur.com/cs544/data/lincoln.txt"
words <- scan(file, what=character())
a) Detect and show all the words that have a punctuation symbol.
b) Replace all the punctuations in the corresponding words with an empty
string. Make this the new words data.
c) Show the frequencies of the word lengths in the above data. Plot the
distribution of these frequencies.
d) What are the words with the longest length?
e) Show all the words that start with the letter p.
f) Show all the words that end with the letter r.
g) Show all the words that start with the letter p and end with the letter r.

```{r}
library(stringr)
file <- "http://kalathur.com/cs544/data/lincoln.txt"
words <- scan(file, what=character())
# a. detect words that contains punct
words[str_detect(words, "[:punct:]")]

# b. Replace punct with empty
words_data <- str_replace(words,"[:punct:]", "" )

# c. get freq of word length + Its distribution
wdata_len <- str_length(words_data)
boxplot(wdata_len, xlab ="word length", ylab = "", horizontal = TRUE )

# d. find the longest words
words_data[which(wdata_len == max(wdata_len))]

# e. words start with p
words_data[str_detect(words_data, "^p")]

# f. words ends with r
words_data[str_detect(words_data, "r$")]

# g. combine e and f
words_data[str_detect(words_data, "^p+.+r$")]

```

##Part2) Data Wrangling (40 points)
Use the tidyverse library for the following:
Download the following csv file,
http://people.bu.edu/kalathur/usa_daily_avg_temps.csv
locally first and use read.csv to load the data into a data frame.
a) Convert the data frame into a tibble and assign it to the variable
usaDailyTemps.
b) What are the maximum temperatures recorded for each year? Show the
values and also the appropriate plot for the results.
c) What are the maximum temperatures recorded for each state? Show the
values and also the appropriate plot for the results.
d) Filter the Boston data and assign it to the variable bostonDailyTemps.
e) What are the average monthly temperatures for Boston? Show the
values and also the appropriate plot for the results.

```{r}
library(dplyr)

temp_data <- read.csv("http://people.bu.edu/kalathur/usa_daily_avg_temps.csv", header = TRUE)
# a.
usaDailyTemps <- as_tibble(temp_data)

# b.
B <- usaDailyTemps %>% group_by(year) %>% summarise(max_temp_by_Year = max(avgtemp))
B
plot(x = B$year, y = B$max_temp_by_Year, type = "l",xlab = "Year", ylab = "Max Avg. Temp recorded")


# c.
C <- usaDailyTemps %>% group_by(state) %>% summarise(maxTemp = max(avgtemp))
C

state_names <- C$state
state_temp <- C$maxTemp


barplot(state_temp,space = 1,
	main = "Max Avg Temp by States", 
	ylab = "temperature(F)",xaxt="n"
	)
labs <- C$state
text(seq(1.5,99.5,by=2), par("usr")[3]-0.25, 
     srt = 60, adj= 1, xpd = TRUE,
     labels = labs, cex=0.65)
##===========================


#6d
bostonDailyTemps <- usaDailyTemps %>% filter(city == "Boston")

#6e
bstByMonth <- bostonDailyTemps %>% group_by(month) %>% summarise(avgMonthlyTemp = mean(avgtemp))
#average boston temperature by month
bstByMonth

plot(x= bstByMonth$month, y = bstByMonth$avgMonthlyTemp, type = "b")

```

