---
title: "CS544_HW4_Vu"
author: "Nguyen Vu"
date: "6/6/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



##Part1) Binomial distribution (20 points)
Suppose a pitcher in Baseball has 50% chance of getting a strike-out when
throwing to a batter. Using the binomial distribution,
a) Compute and plot the probability distribution for striking out the next 6
batters.
b) Plot the CDF for the above
```{r Part 1a, 1b}
#x = strike, prob = 0.5, 
a <- dbinom(0:6, size = 6, prob = 0.5)
plot(0:6, a, type = "h", main = "50% chance strike-out PMF plot", xlab = "# of Strike-out", ylab = "PMF", ylim = c(0, 0.5))
abline(h=0)
points(0:6, a, pch = 16)


aC <- c(0,pbinom(0:6, size = 6, prob = 0.5))
aCDF <- stepfun(0:6, aC)
plot(aCDF, verticals = FALSE, pch = 16, main = "50% chance strike-out CDF plot", xlab = "# of Strike-out", ylab = "CDF", ylim = c(0, 1.1))

```

c) Repeat a) and b) if the pitcher has 70% chance of getting a strike-out.
d) Repeat a) and b) if the pitcher has 30% chance of getting a strike-out.
```{r Part 1c, 1d}
#x = strike, prob = 0.5, 
c <- dbinom(0:6, size = 6, prob = 0.7)
plot(0:6, c, type = "h",main = "70% chance strike out PMF plot", xlab = "# of Strike-outs", ylab = "PMF", ylim = c(0, 0.5))
abline(h=0)

cC <- c(0, pbinom(0:6, size = 6, prob = 0.7))
cCDF <- stepfun(0:6, cC)
plot(cCDF, verticals = FALSE,  pch = 16 ,main = "70% chance strike out CDF plot", xlab = "# of Strike-outs", ylab = "CDF", ylim = c(0, 1.1))


d <- dbinom(0:6, size = 6, prob = 0.3)
plot(0:6, d, type = "h", main = "30% chance strike out PMF plot", xlab = "# of Strike-outs", ylab = "PMF", ylim = c(0, 0.5))
abline(h=0)

dC <- c(0, pbinom(0:6, size = 6, prob = 0.3))
dCDF <- stepfun(0:6, dC)
plot(dCDF, verticals = FALSE, pch = 16, main = "30% chance strike out CDF plot", xlab = "# of Strike-outs", ylab = "CDF", ylim = c(0, 1.1))

```
e) Infer from the shape of the distributions.
The PMF plots shows the probability density of the binomial function base on the chance of striking out. When the chance of striking-out is 50%, the PDF plot resembles a normal distribution. When the chance of striking-out change to 30%, the PDF is right-skewed, and vice versa, when the chance of striking-out increase to 70%, the PDF becomes left skewed. Intuitively, this make sense, if we define the probability of sucess (p) equals strikingout, when p is lower than 50% such as 30%, the probability of "sucess" should be lower at 1, 2 and 3 than at 4, 5 and 6. The reverse is true for when p is higher than 50% such as 70%. When p equals to 50%, we would expect a normal shape distribution as we expect for equal probability of having strikeout a high number of batter and low number of batter. 
Comparing the CDF plots, when p = 50%, the lines for the step plots are move symetrical in step size with the largest steps in the middle. At p = 30% there are more probability that there are less than 4 strike out. The steps are bigger towards the lower values of strike out and much more closer together towards the higher number of strike-outs. The reverse of this is true for p = 70% as the step size are larger towards the higher number of strike-outs. 

##Part 2 Binomial distribution (15 points)
Suppose that 80% of the flights arrive on time. Using the binomial
distribution,
a) What is the probability that four flights will arrive on time in the next 10
flights?
```{r 2a}
# p = 0.8, q = 0.6, find P(x=4), n = 10
dbinom(4 ,size = 10, 0.8)

```
b) What is the probability that four or fewer flights will arrive on time in the
next 10 flights?
```{r 2b}
# find P(x <=4), n = 10
pbinom(4, size = 10, prob = 0.8)


```

c) Compute the probability distribution for flight arriving in time for the next
10 flights.
```{r 2c}
flightpmf <- dbinom(0:10, size = 10, prob = 0.8)
flightpmf

```

d) Show the PMF and the CDF for the next 10 flights.

```{r 2d}
flightpmf <- dbinom(0:10, size = 10, prob = 0.8)
flightcdf <- stepfun( 0:10, c(0, pbinom(0:10, size = 10, 0.8)))

plot(flightpmf, type = "h", main = "PMF of the next 10 flights", xlab = "Flight", ylab = "PMF", pch = 16)
abline(h=0)

plot(flightcdf, verticals = FALSE, main = "CDF of the next 10 flights",xlab = "Flight", ylab = "CDF", pch = 16)
```
## Part3) Poisson distribution (15 points)
Suppose that on average 10 cars drive up to the teller window at your bank
between 3 PM and 4 PM and the random variable has a Poisson
distribution. During this time period,
a) What is the probability of serving exactly 3 cars?
``` {r 3a}
# lambda = 10, x = 3

dpois(3, lambda = 10)

```
b) What is the probability of serving at least 3 cars?

```{r 3b}
ppois(3, lambda = 10)
```
c) What is the probability of serving between 2 and 5 cars (inclusive)?
```{r 3c}
ppois(5, lambda = 10) - ppois(1, lambda = 10)
```
d) Calculate and plot the PMF for the first 20 cars.
```{r 3d}
pmf20c <- dpois(0:20, lambda = 10)

plot(pmf20c, type = "h" ,main = "first 20 cars PMF", xlab = "Car#", ylab = "PMF")
abline(h=0)
```

##Part4) Uniform distribution (15 points)
Suppose that your exams are graded using a uniform distribution between
60 and 100 (both inclusive).
a) What is the probability of scoring i) 60? ii) 80? iii) 100?

```{r 4a}

#because this is a distribution, these three probs will be the same no matter what the score. 
dunif(60, min = 60, max = 100)
dunif(80, min = 60, max = 100)
dunif(100, min = 60, max = 100)
```


b) What is the mean and standard deviation of this distribution?
```{r 4b}
#Mean
sum(c(60:100) * dunif(60:100, min = 60, max = 101))
(60+100)/2

#standard Deviation
sqrt((100-60)^2/12)

```
c) What is the probability of getting a score of at most 70?
```{r 4c}
#P(X <=70)
punif(71, min = 60, max = 100)
sum(dunif(60:70, min = 60, max = 100))


```
d) What is the probability of getting a score greater than 80 (use the
lower.tail option)?
```{r 4d}
punif(80, min = 60, max = 100, lower.tail = FALSE)

```
e) What is the probability of getting a score between 90 and 100 (both
inclusive)?
```{r 4e}
punif(101, min = 60, max = 100) - punif(89, min = 60, max = 100)
```

##Part5) Normal distribution (20 points)
Suppose that visitors at a theme park spend an average of $100 on
souvenirs. Assume that the money spent is normally distributed with a
standard deviation of $10.

a) Show the PDF plot of this distribution covering the three standard
deviations on either side of the mean.
```{r 5a}
threesigma <- 10*3

x <- seq(0, 200)
pkspend_pdf <- dnorm(x, mean = 100, sd = 10)
plot(x, pkspend_pdf, type = 'l', col = 'red', xlim = c(100-threesigma, 100+threesigma), ylim = c(0, 0.05), xaxt = 'n', yaxt = 'n', 
     main = "guests park spending", xlab = "amount", ylab = "freq")
axis(side = 1, at = c(seq(0, 200, by = 10)), labels = TRUE)
axis(side = 2, at = c(seq(0, 0.06, by = 0.01)), labels = TRUE)
```
b) What is the probability that a randomly selected visitor will spend more
than 120?
```{r 5b}
#find P(X>120)

1- pnorm(120,mean = 100, sd=10)

```

c) What is the probability that a randomly selected visitor will spend
between 80 and 90 (inclusive)?
```{r 5c}
pnorm(90,mean = 100, sd=10)- pnorm(80,mean = 100, sd=10)

```
d) What are the probabilities of spending within one standard deviation, two
standard deviations, and three standard deviations, respectively?
```{r 5d}
stdev <- 10
mu <- 100
#prob spending within 1 sd
pnorm(mu+stdev,mean = 100, sd=10)- pnorm(mu-stdev,mean = 100, sd=10)
#prob spending within 2 sd
pnorm(mu+2*stdev,mean = 100, sd=10)- pnorm(mu-2*stdev,mean = 100, sd=10)
#prob spending within 3 sd
pnorm(mu+3*stdev,mean = 100, sd=10)- pnorm(mu-3*stdev,mean = 100, sd=10)
```
e) Between what two values will the middle 90% of the money spent will
fall?
```{r 5e}
#upper limit
qnorm(0.95, mean = 100, sd = 10)
#lower limit
qnorm(0.05, mean = 100, sd = 10)
```
f) Show a plot for 10,000 visitors using the above distribution.
```{r 5f}
#simulate 10000 guest spending
f <- rnorm(10000, mean = 100, sd = 10) 
plot(table(round(f)), type = "h", main = "Spending of 10K park guests",ylab = "Frequency", xlab = "Spending")
abline(h=0)

```


## Part6) Exponential distribution (15 points)
Suppose your cell phone provider’s customer support receives calls at the
rate of 18 per hour.

rate = 18/hr

a) What is the probability that the next call will arrive within 2 minutes?
```{r 6a}
# 2min = 2/60 hour
pexp(2/60, rate = 18)

```

b) What is the probability that the next call will arrive within 5 minutes?
```{r 6b}
# 5min = 5/60 hour
pexp(5/60, rate = 18)

```
c) What is the probability that the next call will arrive between 2 minutes
and 5 minutes (both inclusive)?
```{r 6c}
# 2min = 2/60 hour
pexp(5/60, rate = 18) - pexp(2/60, rate = 18)

```
d) Show the CDF of this distribution.
```{r 6d}
x <- seq(0, 1, by=1/60)
cdf <- pexp(x, rate = 18)

plot(x, cdf, type = "l", col= "blue", xlim = c(0, 0.5), main = "Cummilative prob of calls arriving", 
     xlab = "hours", ylab = "Probability")

```

