---
title: "Practice Finals_2018_Vu"
author: "Nguyen Vu"
date: "6/17/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Packages
```{r cars}
library(dplyr)
library(prob)
library(sampling)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

##Q1. (30 points)
a)
Given a vector of numeric values, write a single simplest R expression that returns TRUE if all
the odd index elements of the vector are twice their succeeding value, otherwise your
expression returns FALSE. You can assume that the given vector has an even number of
values and is assigned to a variable named data.
TRUE scenario:
c(6, 3, 10, 5, 22, 11, 26, 13)
FALSE scenario:
c(7, 3, 6, 5, 22, 11, 26, 13)

```{r}
test_true <- c(6, 3, 10, 5, 22, 11, 26, 13)

test_false <- c(7, 3, 6, 5, 22, 11, 26, 13)

all(test_true[seq(1, length(test_true), 2)]/test_true[seq(1, length(test_true), 2)+1] == 2)

all(test_false[seq(1, length(test_false), 2)]/test_false[seq(1, length(test_false), 2)+1] == 2)

```
b)
Using a for loop, write the R function, evensum(data), that returns the
sum of all the even values in the given numeric vector.
Now, without using any loop, write the R function, evensum2(data), that returns the
sum of all the even values in the given numeric vector.
Test both functions with sample data
c(0,2,3,4,7,8,6,9)

```{r}
data <- seq(1,99)

evensum <- function(vector){
  dftsum = 0
  for (i in vector){
    if(i%%2 == 0 ){
      dftsum = dftsum + i
    }
  }
  return(dftsum)
}

evensum(data)

evensum2 <- function(vector){
  vdf<-data.frame(
    vec = c(vector),
    checkEven = c(vector%%2)
  )
  totals = sum(subset(vdf, vdf$checkEven !=1)$vec)
  return(totals)
}

evensum2(data)
```
c)
Given a vector, show the R code for finding the element in the vector has the longest run (same
value consecutively)?
Sample Input:
c(10,20,20,40,40,40,40,40,50,50,10,40)
Output:
40

```{r}

x <- c(10,20,20,40,40,40,40,40,50,50,10,40)

x[max(rle(c(x))$lengths)]

```

##Q2. (10 Points)
Consider a class of students with 6 boys and 4 girls. Setup the experiment for selecting 5
unique students for representing the class as one group at the science fair. Add the probability
space for this experiment. Add the random variable for the number of boys. Show the marginal
distribution for this random variable. Add the random variable for the number of girls. Show the
marginal distribution for this random variable. Provide the R code for all the above.

```{r}
classroom <- urnsamples(c(rep("boy", 6), rep("girl", 4)), size = 5, replace = FALSE)

S<-probspace(classroom)

Prob(S, isrep(S, "boy", 2))

S <- addrv(S, N_boy = rowSums(S == "boy"))

marginal(S, vars = "N_boy")

```

Q3. (15 Points)
Consider the Titanic dataset from R.
Show how you would compute the following using simplest R code. Provide the R code and the
results for all the following.
a) How many males and females are there in the data set?
b) By gender, how many survived and did not survive?
c) By traveling class, how many survived and did not survive?
d) By traveling class and age, how many survived and did not survive?

```{r}
data("Titanic")
Titanic

#a
apply(Titanic, c(2), sum)
#b
apply(Titanic, c(2, 4), sum)
#c
apply(Titanic, c(1, 4), sum)
#d
apply(Titanic, c(1, 3, 4), sum)

```

Q4. (15 points)
A random variable has normal distribution mean 100 and standard deviation 5. Show
how you would do the following with R code. Write the solutions and interpretation in the
box below.
a) Determine and interpret the quartiles of this variable.
b) Find the value that 65% of all possible values of this variable exceed.
c) Find the two values that divide the area under the corresponding normal curve into a
middle area of 0.95 and two outside areas of 0.025. Interpret your answer

```{r}

ranVar <- rnorm(1, mean = 100, sd = 5)

pmf <- dnorm(c(1:200), mean = 100, sd = 5)
quantile(pmf)

plot(pmf, type = "l")
#b
qnorm(0.65, mean = 100, sd = 5)

#c
midArea = 0.95
tailArea = .025

qnorm(tailArea, mean = 100, sd = 5)
qnorm(midArea+tailArea, mean = 100, sd = 5)
# this is bound of the 2 sided 95% confidence interval around the mean = 100

```
##Q5. (15 points)
Consider the votes.repub dataset from “cluster” library, initialized with the following
code. Show how you would do the following with R code.
library(cluster)
library(sampling)
data(votes.repub)
data <- votes.repub
The rownames of the data frame are the names of the states.
a) Add the column named Letter to the data frame initialized with the first letter of the
state. Use substring(rownames(data), 0, 1) for this purpose.
b) Show the contingency table for this Letter.
c) Using the stratified sampling with the Letter as the stratum, select a sample that
selects one state from each stratum.
d) Compare the means of the data for the year 1976 with the mean of the selected
sample
```{r}
#Load data n packages
library(cluster)
library(sampling)
data(votes.repub)
data <- votes.repub

#b
data$Letter <- substring(rownames(data), 0, 1)
#contingency table
table(data$Letter)

#c
#strata size length
sts_len <- length(table(data$Letter))
#stra size to sample ONE state from each letter
sts <- rep(1, sts_len)
#select strata
strtSamp <- strata(data, stratanames = c("Letter"), size = sts, method = "srswor")

st_Samples <- getdata(data, strtSamp)

#d
#exclude NA cells... Approach 1 replace NA with O to keep data frame shape...?
pop1976 <- mean(data$X1976[!is.na(data$X1976)])
st1976 <- mean(st_Samples$X1976[!is.na(st_Samples$X1976)])
cat("\n","The mean for the year 1976 in the sample population is ", pop1976, "\n", "The mean for the year 1976 in the selected strata sample is ", st1976)

```

##Q6) (15 points)
Use the stringr library for the following. Given an input string, write the simplest and
elegant R code without using explicit loops for the following.
a) a vector of all prefixes of the input string. (e.g., for input string “abc”, the prefixes are
“”, “a”, “ab”, “abc”)
b) a vector of all suffixes of the input string. (e.g., for input string “abc”, the suffixes are
“abc”, “bc”, “c”, “”)
c) Use the fruit data from this library. Which starting letter has the highest number of
occurrences in this data? Which ending letter has the highest number of occurrences in
this data? If there is a tie, use the first result. Provide R code and results.
d) Using the results from c), show the fruits that have that starting letter and that ending letter.
Provide R code and results.

```{r}
library(stringr)
x_str <- "ThisIsAtestString"

#a
str_sub(x_str, start = c(0) ,end = c(0:str_length(x_str)))
#b
str_sub(x_str, start = c(1:str_length(x_str)), end = c(-1))

#C
data("fruit")
# Which starting letter has the highest number of
# occurrences in this data? 
start_letters_Freq <- str_count(str_c(substr(fruit, 1, 1), collapse = ''), letters)
max_st_letter <- letters[which(start_letters_Freq == max(start_letters_Freq))]
#Which ending letter has the highest number of occurrences in
# this data? 
end_letters_Freq <- str_count(str_c(substr(fruit, str_length(fruit), str_length(fruit)), collapse = ''), letters)
max_end_letter <- letters[which(end_letters_Freq == max(end_letters_Freq))]

#d
#show fruits with starts with "p" and end with "e" and "t"

pat <-"^[p]\\w*[e,t]$"
fruit[str_detect(fruit, pattern = pat)]
```