---
title: "CS555_term_project_Vu"
author: "Nguyen Vu"
date: "8/21/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load packages}
library(dplyr)
library(stats)
library(car)
library(tcltk)
library(asbio)
library(moments)
library(ggplot2)
library(ggcorrplot)
```

## Load Data File and Data Clean Up and Adding Variables  


#### Location of data on Kaggle
#### https://www.kaggle.com/goldenoakresearch/us-acs-mortgage-equity-loans-rent-statistics

```{r Load Data File and Data Clean Up and Adding Variables}
#load Data File

#colums of interests:
#c(1, 6, 7, 14:18, 22, 25, 37, 40, 47, 50, 52, 54)
#read in giant CSV file (80 variables and 39030 observations):
masterDT <- read.csv("real_estate_db.csv", header = T)
# verify the names of the columns of interests above before subsetting:
vars_names <- names(masterDT[,c(1, 6, 7, 14:18, 22, 37, 47, 52)])
vars_names
#subset columns that contains columns of interests and replace the original df with the smaller df
masterDT <- masterDT[,c(1, 6, 7, 14:18, 22, 37, 47, 52)]#now data is reduced to 12 variables and 39030 observations
#remove rows of incomplete cases
masterDT <- masterDT[complete.cases(masterDT),] #data is reduced to 12 variables and 37942 observations
#create a back up of the data frame incase there is an error during analysis. 
masterDT_backup <- masterDT

#rename columns to be more readable:
new_vars_names <- gsub("hc", "Owner_Costs", vars_names)
new_vars_names <- gsub("hi_", "Household_Income_", new_vars_names)
names(masterDT) <- new_vars_names

#Create a varible to define the continental US vs states like Hawaii, Alaska and Puerto Rico to accurately
#study at the effect longitude and latitude
# Continental state = 1, non-continental states = 0
masterDT$Continental <- ifelse(masterDT$state_ab %in% c("HI", "AK", "PR"), 0, 1)

# Create a value to define the the ratio of water to land ratio of specific geographic area
masterDT$water_land_ratio <- masterDT$AWater/masterDT$ALand

#creat a subset of just the continental US for analysis for Hawaii, Alaska and Puertorico can be considered geographically outliers
contUSdf <- filter(masterDT, Continental !=0) #18 variables and 37228 observations
attach(contUSdf)#attach to memory for ease of analysis

# Let longitude of the US be devided into 3 equal segments.
#Westcoast will be defined as lowest 1/3rd segment, Eastcoast will be the higest 1/3rd segment and Central will be the middle segment
lng_seg<- (max(contUSdf$lng)-min(contUSdf$lng))/3
contUSdf$Coast <- cut(contUSdf$lng, breaks = c(-Inf, min(contUSdf$lng)+lng_seg,  min(contUSdf$lng)+2*lng_seg, Inf  ),
                      labels = c("West", "Middle", "East"))

#let latitude of the US be divided into 2 equal segments,
# North will be defined as the highest half and the south will be defined as the lowest half.
contUSdf$Region <- cut(contUSdf$lat, breaks = c(-Inf, mean(contUSdf$lat), Inf), 
                       labels = c("South", "North"))


# Calculate for the Population Density by divigin population by land Area
contUSdf$Pop_density <- contUSdf$pop/contUSdf$ALand
```


## Simple look at the data  

Get the top 3 and bottom 3 states in term of highest and lowest gross rental median and owner's total costs including mortgage

Get the average median rent, owner's costs including mortgage


```{r simple analysis}
#Which State has the lowest average median cost for owner interm of Mortgage and cost?
s<- contUSdf %>% group_by(state) %>% summarise(avg_mortgage_plus_cost = mean(Owner_Costs_mortgage_median), 
                                               avg_med_rent = mean(rent_median))
s %>% arrange(avg_mortgage_plus_cost)%>% slice(1:3)

#which State has the highest average median cost for owner of Mortgage and cost?
s %>% arrange(desc(avg_mortgage_plus_cost))%>% slice(1:3)

#Which state has the lowest average median gross median rent?
s %>% arrange(avg_med_rent)%>% slice(1:3)

#Which state has the highest average median gross median rent?
s %>% arrange(desc(avg_med_rent))%>% slice(1:3)

# get a table of averages summary of the average of Rent, Owner Cost, Owner Costs Mortgage and household Income by Coasts and region
contUSdf%>% group_by(Coast, Region) %>% summarise(US_avg_median_rent = mean(rent_median),
                                                               US_avg_median_Owner_Costs = mean(Owner_Costs_median),
                                                               US_avg_median_Owner_Costs_n_Mortgage = mean(Owner_Costs_mortgage_median),
                                                               US_avg_median_householdIncome = mean(Household_Income_median))
```


## More analysis by Location...  
From the table above, it looks like the average median rent is the roughly the similar between regions (north vs south) but may be different between coasts (East, Middle and West). Check for this claim at the 95%CI

```{r Rent by Location category}
attach(contUSdf)
p <- ggplot(data = contUSdf) #load data in a plot
# Graph rent distribution by boxplot
p + geom_boxplot(aes(x = Coast, y = rent_median)) + facet_grid(Region~.) + xlab("Coast") + ylab("Median Rent")

#### Gross Rent Analysis by Location: East VS West VS Middle coasts and North VS South region
#Compare the average median rent the Northern VS. Sourthern US, and East Coast Vs. West Coast VS. Middle at the 95% CI:

# Global F test looking at only Region:
summary(aov(rent_median~Region)) #not significant in global F test.
p + geom_boxplot(aes(x = Region, y = rent_median)) + ylab("Median Rent") #Region ONly Plot

#Compare the average median rent by coasts alone
summary(aov(rent_median~Coast)) #Global F Test is significant
p + geom_boxplot(aes(x = Coast, y = rent_median)) #Coast Only Plot
pairwise.t.test(rent_median,Coast) #differences accross East Coasts, West Coast and Middle US. 

```
No difference when comparing regions.   
F test indicate there is a difference between coasts and pairwise t test confirms that there is a significant diffence between the 3 coasts: East, West and Middle. 
Perform Ancova to check for covariance when fitting region and coasts. Check for interaction and fit model. 


```{r by location category cont}
#Fit Rent when control for both Coasts and Region
m_test<-lm(rent_median~Coast*Region)
Anova(m_test, type = 3) #There is Interactions effect accross Groups
#box plot of Coasts with Region by Color
p + geom_boxplot(aes(x = Region, y = rent_median, colour = Coast))
#split data up by Region to stratify data to control for interaction

```

Interaction between coasts and regions exists. Stratifying is needed. Control for Region, test for difference between coast. Use West as the reference group. 

```{r by location category cont2}
m_north <- lm(rent_median[Region=="North"] ~ Coast[Region == "North"])
m_south <- lm(rent_median[Region=="South"] ~ Coast[Region == "South"])
Anova(m_north, type = 3)#F significant
Anova(m_south, type = 3)#F Significant
#West is the reference group
fitted_m_north <- fitted(m_north)
summary(m_north) #in the north, no different in rent between West and East
# In the north, there is a difference between Middle and West 

fitted_m_south <- fitted(m_south)
summary(m_south) # in the south, rent is difference between middle and east and west
```


## Multivariate Regression Analysis of Median Gross Rent...  
### Correlations
It would be intersting to look at the other variables available to see if we can estimate any other factor that influence Rent
```{r Rent model}
attach(contUSdf)
# Graphicsal Correlations analysis of variables
contUSdfFactors <- data.frame(Pop_density, water_land_ratio, Household_Income_median, rent_median, Owner_Costs_mortgage_median, Owner_Costs_median)

#Check for correlation
cor(contUSdfFactors)
c <- round(cor(contUSdfFactors), 2)
ggcorrplot(c, type = "lower", lab = T)
```

Strong correlation:  
* House hold income and Rent  
* House hold income and Owner Cost + Mortgage  
* House hold income and Owner Cost  
* Owner Cost + Mortgage and Owner Cost  

Moderate correlation for Owner Cost and Rent
  
Weak or no correlation between Population Density, and water:land ratio

### MlR analysis of median rent

```{r MLR}
# MLR for Rental median

#Stratify to control for north south region because of interaction found in the previous section
contUSdf_N <- contUSdf[Region=="North",]
contUSdf_S <- contUSdf[Region=="South",]
# Create dummy Variables for Coasts and Region for multiple regression analysis
contUSdf_N <- contUSdf %>% mutate(
  gWest = if_else(Coast=="West", 1, 0),
  gMid = if_else(Coast=="Middle", 1, 0),
  gEast = if_else(Coast=="East", 1, 0)
)
attach(contUSdf_N)
rent_mlrm <- lm(rent_median ~ Household_Income_median+
                  Pop_density+water_land_ratio+
                  Owner_Costs_mortgage_median+gWest+gMid)
anova(rent_mlrm) #F ratio is significant
summary(rent_mlrm)
#R Square = 0.6367 
suma <- summary(rent_mlrm)
#get the slope estimates table for each predictor
suma$coefficients
#get 95% interval of slope estimates for each predictor
confint(rent_mlrm)

contUSdf_N$fitted_rent <- unclass(fitted(rent_mlrm))
contUSdf_N$resid_rent <- unclass(resid(rent_mlrm))

hist(contUSdf_N$resid_rent, xlab = "Residual Gross Rent")
skewness(contUSdf_N$resid_rent) #only slightly skewwed but resembled normal at skewness = 0.58

ggplot(data = contUSdf_N) + geom_point(aes(x=fitted_rent, y=resid_rent))+ geom_smooth(aes(x=fitted_rent, y=resid_rent), method = "lm", se = T) + xlab("North Region Predicted Rent")+ ylab("Residual")
#



# Create dummy Variables for Coasts and Region for multiple regression analysis
contUSdf_S <- contUSdf %>% mutate(
  gWest = if_else(Coast=="West", 1, 0),
  gMid = if_else(Coast=="Middle", 1, 0),
  gEast = if_else(Coast=="East", 1, 0)
)
attach(contUSdf_S)
rent_mlrm2 <- lm(rent_median ~ Household_Income_median+
                  Pop_density+water_land_ratio+
                  Owner_Costs_mortgage_median+gWest+gMid)

summary(rent_mlrm2)
anova(rent_mlrm2)
suma2 <- summary(rent_mlrm2)
#get the slope estimates table for each predictor
suma2$coefficients
#get 95% interval of slope estimates for each predictor
confint(rent_mlrm2)

#get predicted and residuals
contUSdf_S$fitted_rent <- unclass(fitted(rent_mlrm2))
contUSdf_S$resid_rent <- unclass(resid(rent_mlrm2))
attach(contUSdf_S)
hist(contUSdf_N$resid_rent, xlab = "South Residual Gross Rent")
skewness(contUSdf_N$resid_rent) #only slightly skewwed but resembled normal at skewness = 0.62

ggplot(data = contUSdf_S) + geom_point(aes(x=fitted_rent, y=resid_rent))+ geom_smooth(aes(x=fitted_rent, y=resid_rent), method = "lm", se = T) + xlab("South Predicted Rent")+ ylab("Residual")


```

In both region, South and North, we've found evidence that rental prices is in fluenced by all tested factors.  

## Proportion test of pricy rental above 90 percentile of median rent  

```{r Proportion of pricy rentals}
##############################
##############################
# Comparing the states with the highest proportion of rent above 75 percentile price. 
#get the 90 percentile of all the median rent in the US
US90PercentileRent <- quantile(rent_median, 0.90)

# Create variable Top10% of area with median above the 90 percentile of the national median
# 1 is above, 0 is below
contUSdf$top10percent <- if_else(rent_median>US90PercentileRent, 1, 0)

#get the proportion of homes by coasts with proportion of median rent in the top 10% of the nation
contUSdf %>% group_by(Coast) %>% count(top10percent==1)
contUSdf %>% group_by(Coast) %>% summarise(n = n())
#Test the hypotheis that the East Coast has just as % rental market with median prices higher than 
# 90% of the median as the west coasts

prop.test(c(1789,1465), c(18239,7491), conf.level = 0.95)
#The % of rental in the East Coast to be outrageous pricing above the 90% of national median is actually less than the west by
# 8.7% to 10%. 
```
The % of rental in the East Coast to be outrageous pricing above the 90% of national median is actually less than the west by
# 8.7% to 10%. 